\propertitle

### 2019-11-19 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-19T14:47:34+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

These two steps are more straightforward if instead of power moments we use
\emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
factorial moment of a distribution $\pxx(\ya)$ for the activity of the
sample neurons is defined as the average
\begin{equation}
  % \E\condt[\Bigg]{
  %   \binom{\ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  % \equiv
  \sum_{\ya=0}^{n} %m!\,
  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
  \qquad 1\le m \le n
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
This moment can be interpreted as the expectation of the number of distinct
$m$-tuples of simultaneously spiking neurons (within a bin's time width),
normalized by the number of distinct $m$-tuples. For example, with $m=2$,
if $a=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
distinct pairs of simultaneously spiking neurons, and the total number of
distinct pairs is $\tbinom{5}{2}=10$. The normalized number of spiking
pairs is therefore $6/10$. Note that the first $m$ factorial moments
provide the same information as the first $m$ power moments and vice versa:
they linearly determine each other because $\tbinom{\ya}{m}$ is a
polynomial in $\ya$ of degree $m$. Specifying one set is therefore
equivalent to specifying the other set. But the normalized factorial
moments have a convenient property for our analysis
\citep[\sect~2.3]{portamanaetal2015}{potts1953}: \emph{the first $n$
  normalized factorial moments for the sample and for the full network are
  numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
    \sum_{\ya=0}^{n} %m!\,
  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
=    \sum_{\yA=0}^{N} %m!\,
  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),
  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad 1\le m \le n,
\end{equation}
where $\pxxx(\yA)$ is our distribution of belief about the full-network
activity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Appendices
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%\clearpage
% %\renewcommand*{\appendixpagename}{Appendix}
% %\renewcommand*{\appendixname}{Appendix}
% %\appendixpage
% \appendix

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-19T16:21:29+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

These two steps are more straightforward if instead of power moments we use
\emph{normalized factorial moments} \citep{potts1953}. The $m$th normalized
factorial moment of a distribution $\pxx(\ya)$ for the activity of the
sample neurons is defined as the average
\begin{equation}
  \label{eq:def_factorial_moment_a}
  % \E\condt[\Bigg]{
  %   \binom{\ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  % \equiv
  \sum_{\ya=0}^{n} %m!\,
  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
  \qquad 1\le m \le n
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
This moment can be interpreted as the expectation of the number of distinct
$m$-tuples of simultaneously spiking neurons (within a bin's time width),
normalized by the number of distinct $m$-tuples. For example, with $m=2$,
if $a=4$ neurons spike in a network of $n=5$, we have $\tbinom{4}{2}=6$
distinct pairs of simultaneously spiking neurons, and the total number of
distinct pairs is $\tbinom{5}{2}=10$. The normalized number of spiking
pairs is therefore $6/10$. Note that the first $m$ factorial moments
provide the same information as the first $m$ power moments and vice versa:
they linearly determine each other because $\tbinom{\ya}{m}$ is a
polynomial in $\ya$ of degree $m$. Specifying one set is therefore
equivalent to specifying the other set. But the normalized factorial
moments have a convenient property for our analysis
\citep[\sect~2.3]{portamanaetal2015}{potts1953}: \emph{the first $n$
  normalized factorial moments for the sample and for the full network are
  numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
    \sum_{\ya=0}^{n} %m!\,
  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
=    \sum_{\yA=0}^{N} %m!\,
  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),
  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad 1\le m \le n,
\end{equation}
where $\pxxx(\yA)$ is our distribution of belief about the full-network
activity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Appendices
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%\clearpage
% %\renewcommand*{\appendixpagename}{Appendix}
% %\renewcommand*{\appendixname}{Appendix}
% %\appendixpage
\appendix

\section{Derivation of the maximum-entropy distribution}
\label{sec:derivation_maxent_fullnet}

Here is a summary derivation of the maximum-entropy distribution for the
full network constrained by a set of factorial moments. For further details
see Porta Mana \etal\ \citey{portamanaetal2015}.

First of all the sampling relation. We have a set of $N$ units, $\yA$ of
which have some specific property, and we sample in an unknown way $n$ of
the $N$ units. The probability that $\ya$ of the $n$ sampled units have that
property is then given by the hypergeometric distribution
\begin{equation}
  \label{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{n\ya}\binom{\yNv-n}{\yNv \yA-n\ya}\binom{\yNv}{\yNv \yA}^{-1}.
\end{equation}
typical of \enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In the following leave $n$, $N$ implicit in the conditional. If we are
uncertain about the number $\yA$, with belief $P(\yA)$, then by the theorem
of total probability our belief about $\ya$ is
\begin{equation}
  \label{eq:belief_sample_unknown_A}
  p(\ya)= \sum_{\yA} \yG_{\ya\yA} \;P(\A).
\end{equation}
In our case the units are neurons, the set is the full network, and the
property is their being active in a specific time bin.

From the definition of normalized factorial
moment~\eqref{eq:def_factorial_moment_a}, the expression for the hypergeometric
distribution~\eqref{eq:conditional_hypergeometric}, and the
relation~\eqref{eq:belief_sample_unknown_A} between our beliefs about $\ya$
and $\yA$, it is possible to prove using some combinatorial juggling the
equality~\eqref{eq:equal_m_factorialmoments} between the factorial moments
of sample and full network.

For the construction of a maximum-entropy distribution from generic
constraints see Jaynes \citey{jaynes1963}[\chap~11]{jaynes1994_r2003}. More
precisely we use the minimum relative-entropy principle
\citep{hobsonetal1973,csiszar1985}[\sect~5.2.2]{sivia1996_r2006} with
respect to a reference distribution. For the solution of the extremization
problem see Mead \etal\ \citey{meadetal1984} and the extensive text by Fang
\etal\ \citey{fangetal1997}.

\section{Measure of informational sufficiency}
\label{sec:measure_suff}
